Robust Estimation, Prediction and Control with Dynamics Structure

### 2020-01-13 main.tex
Updates from Overleaf

Differences with our work:
\begin{itemize}
    \item We with to act in a reasonable (structure) but cautious (robust) way, and progressively get more and more aggressive as we obtain more data and reduce uncertainty.
    \item Thus, we do not wish to explore actively the more promising regions as in OFU, nor the more likely region as in TS, nor act randomly as in Coarse-ID.
    \item We have linear dynamics but our cost is not quadratic (to accurately represent planning task with discrete/branching decisions: e.g. take or give way)
    \item Our optimisation problem amounts to discrete controller selection, instead of continuous controller synthesis

@@

\begin{lemma}[Confidence polytope]
\label{lem:confidence_polytope}
We can enclose the confidence ellipsoid $C_\delta$ obtained in $\eqref{eq:confidence-ellipsoid}$ within a polytope $P_\delta$:
\[\cC_\delta \subset \cP_\delta\]
where $\cP_\delta$ can be written as:
\begin{equation}
     \cP_\delta = \left\{ A_{0}+\sum_{i=1}^{2^d}\lambda_{i}\Delta A_{i}: \lambda\in[0, 1]^{2^d},  \sum_{i=1}^{2^d}\lambda_{i}=1\right\}.
\end{equation}
with 
\begin{align*}
    &G_{N,\lambda}^{-1} = PDP^{-1}\\
    &h_k \text{ is the }k^\text{th}\text{ element of }\{-1,1\}^d\text{, for } k\in[2^d]\\
    &\Delta\theta_k = C_{N,\lambda,\delta}^{1/2} P^{-1}D^{-1/2} h_k \\
    &A_0 = \theta_{N,\lambda}^\transp\Phi\\
    &\Delta A_k = \Delta\theta_k^\transp\Phi\\
\end{align*}
\end{lemma}

\begin{figure}
    \centering
    \includegraphics[trim={3.8cm, 2cm, 5cm, 3.8cm}, clip, width=0.6\textwidth]{img/ellipsoid_to_polytope}
    \caption{From the confidence ellipsoid $\cC_\delta$ to its enclosing polytope $\cP_\delta$}
    \label{fig:ellipsoid_to_polytope}
\end{figure}

@@

Then, 
\[
\|\theta_{N,\lambda} - \theta\|_1 \leq \|G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\|_1\|\eta\|_\infty + \lambda\|G_{N, \lambda}^{-1}\Sigma_d\theta\|_1
\]

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{img/interval-hull}
    \caption{A few trajectories are sampled from an initial state $x_0$ following a policy $\pi=(u_0,\cdots)$ with various parameters $\theta$ (in black). The union of reachability sets is shown in green, and its interval hull in red.}
    \label{fig:interval-hull}
\end{figure}

@@

\begin{algorithm}[tp]
  \SetAlgoLined\DontPrintSemicolon
  \SetKwFunction{algo}{robust\_control}
  \SetKwFunction{proc}{pessimistic\_reward}
  \SetKwProg{myalg}{Algorithm}{}{}
  \myalg{\algo{$s_0$}}{
  \While{resources available}
  {
  Generate a candidate sequence $\pi$ with a planning algorithm using \proc{} instead of the true rewards.
  }
  \KwRet $\argmax_{\pi\in\Pi} \hat{v^r}(\pi)$ \;}{}
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\proc{$s_0$, $\pi$, T}}{
  Integrate \eqref{eq:interval_predictor} over $[0,t]$ with commands $\pi$ to compute $[\underline{x}(t), \overline{x}(t)]$\;
  \KwRet $\min_{x\in[\underline{x}(t), \overline{x}(t)]} r(x)$\;}
\caption{Interval-based Robust Control}
\label{algo:irc}

@@

\begin{proof}
Let $\theta\in\Theta$, by \autoref{thm:confidence_interval} we have that $A(\theta) \in \cP$ with probability $1-\delta$. Whenever it is the case, we have by \autoref{thm:predictor} that the inclusion property \eqref{eq:interval_property} is verified by the interval predictor \eqref{eq:interval_predictor} for all $\omega(t)\in[\underline{\omega}(t), \overline{\omega}(t)]$. This gives the following bound:
\begin{equation*}
 \sum_{t=0}^\infty \gamma^t r(x_t) \geq \sum_{t=0}^\infty \min_{x\in[\underline{x}(t, \pi), \overline{x}(t, \pi)]} \gamma^t r(x) = \hat{v^r}(\pi)
\end{equation*}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{img/robust-control-tree}
\caption{The computation of robust b-values in Algorithm \ref{algo:drop}. The simulation of trajectories for every dynamics model $f^m$ is represented as stacked versions of the expanded tree $\mathcal{T}_n$.}
\label{fig:drop}
\end{figure}

@@

\begin{algorithm}[tp]
\DontPrintSemicolon
Initialize $\mathcal{T}$ to a root and expand it. Set $n=1$.\;
\While{Numerical resource available}{
Compute the robust u-values $u^r_i(n)$ and robust b-values $b^r_i(n)$.\;
Expand $\argmax_{i\in \mathcal{L}_n} b^r_i(n)$.\;
n = n + 1\;
}
\Return $\argmax_{a\in \mathcal{A}} u^r_a(n)$
\caption{Deterministic Robust Optimistic Planning}
\label{algo:drop}
\end{algorithm}

@@

>>>>

Differences with our work:
\begin{itemize}
    \item We wish to act straight away in a reasonable (structure) but cautious (robust) way, and progressively get more and more aggressive as we obtain more data and reduce uncertainty.
    \item Thus, we do not wish to explore actively the more promising regions as in OFU, nor the more likely region as in TS, nor act randomly as in Coarse-ID.
    \item We have linear dynamics but our cost is not quadratic. Thus, we can accurately represent planning tasks with discrete/branching decisions involving discontinuities (e.g. collision vs non-collision states).
    \item Our optimisation problem amounts to a sequence of controllers selection, instead of continuous controller synthesis

@@

\begin{lemma}[Confidence polytope]
\label{lem:confidence_polytope}
We can enclose the confidence ellipsoid $C_\delta$ obtained in $\eqref{eq:confidence-ellipsoid}$ within a polytope $P_\delta$:
\[\cC_\delta \subset \cP_\delta\]
where $\cP_\delta$ can be written as:
\begin{equation}
     \cP_\delta = \left\{ A_{0}+\sum_{i=1}^{2^d}\lambda_{i}\Delta A_{i}: \lambda\in[0, 1]^{2^d},  \sum_{i=1}^{2^d}\lambda_{i}=1\right\}.
\end{equation}
with 
\begin{align*}
    &G_{N,\lambda}^{-1} = PDP^{-1}\\
    &h_k \text{ is the }k^\text{th}\text{ element of }\{-1,1\}^d\text{, for } k\in[2^d]\\
    &\Delta\theta_k = C_{N,\lambda,\delta}^{1/2} P^{-1}D^{-1/2} h_k \\
    &A_0 = \theta_{N,\lambda}^\transp\Phi\\
    &\Delta A_k = \Delta\theta_k^\transp\Phi\\
\end{align*}
This conversion is illustrated in \autoref{fig:ellipsoid_to_polytope}
\end{lemma}

\begin{figure}
    \centering
    \includegraphics[trim={3.8cm, 2cm, 5cm, 3.8cm}, clip, width=0.4\textwidth]{img/ellipsoid_to_polytope}
    \caption{From the confidence ellipsoid $\cC_\delta$ to its enclosing polytope $\cP_\delta$}
    \label{fig:ellipsoid_to_polytope}
\end{figure}

@@

Then, 
\begin{align*}
\|\theta_{N,\lambda} - \theta\|_1 &\leq \|G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\|_1\|\eta\|_\infty + \lambda\|G_{N, \lambda}^{-1}\Sigma_d\theta\|_1\\
&\leq \|G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\|_1\|\eta\|_\infty + \lambda\|G_{N, \lambda}^{-1}\Sigma_d\|_1\|\theta\|_\infty
\end{align*}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.3\textwidth]{img/interval-hull}
    \caption{A few trajectories are sampled from an initial state $x_0$ following a policy $\pi=(u_0,\cdots)$ with various parameters $\theta$ (in black). The union of reachability sets is shown in green, and its interval hull in red.}
    \label{fig:interval-hull}
\end{figure}

@@

\begin{algorithm}[tp]
  \SetAlgoLined\DontPrintSemicolon
  \SetKwFunction{algo}{robust\_control}
  \SetKwFunction{proc}{pessimistic\_step}
  \SetKwProg{myalg}{Algorithm}{}{}
  \myalg{\algo{$s_0$}}{
  \While{resources available}
  {
  Select a candidate action sequence $a$ with a planning algorithm.\;
  Sample a pessimistic transition and reward $\underline{r}_t$ using \proc{$\underline{x}(t)$, $\overline{x}(t)$, a}.
 \;
 Update the planning algorithm with the sampled pessimistic transition.
  }
  \KwRet $\argmax_{\pi\in\Pi} \hat{v^r}(\pi)$ \;}{}
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\proc{$\underline{x}(t)$, $\overline{x}(t)$, $a$}}{
  Integrate \eqref{eq:interval_predictor} over $dt$ with controls $a$ to compute $\underline{x}(t+dt), \overline{x}(t+dt)$\;
  Compute worst-case reward $\underline{r}_t = \min_{x\in[\underline{x}(t), \overline{x}(t)]} r(x)$\;
  \KwRet $\underline{x}(t+dt), \overline{x}(t+dt), \underline{r}_t$\;}
\caption{Interval-based Robust Control}
\label{algo:irc}

@@

@@

\begin{figure}
\centering
\includegraphics[width=0.3\linewidth]{img/robust-control-tree}
\caption{The computation of robust b-values in Algorithm \ref{algo:drop}. The simulation of trajectories for every dynamics model $f^m$ is represented as stacked versions of the expanded tree $\mathcal{T}_n$.}
\label{fig:drop}
\end{figure}

@@

\begin{algorithm}[tp]
\DontPrintSemicolon
Initialize $\mathcal{T}$ to a root and expand it. Set $n=1$.\;
\While{Numerical resource available for episode $n$}{
Compute the robust b-values $b^r_i(n)$.\;
Expand the leaf $\argmax_{i\in \mathcal{L}_n} b^r_i(n)$.\;
}
\Return $\argmax_{a\in \mathcal{A}} u^r_a(n)$
\caption{Deterministic Robust Optimistic Planning}
\label{algo:drop}
\end{algorithm}

@@

\subsection{Proof of \autoref{prop:lower-bound}}

\begin{proof}
We have that $A(\theta) \in \cP$ with probability $1-\delta$. Whenever it is the case, we have by \autoref{thm:predictor} that the inclusion property \eqref{eq:interval_property} is verified by the interval predictor \eqref{eq:interval_predictor} for all $\omega(t)\in[\underline{\omega}(t), \overline{\omega}(t)]$. This gives the following bound:
\begin{equation*}
 \sum_{t=0}^\infty \gamma^t r(x_t) \geq \sum_{t=0}^\infty \min_{x\in[\underline{x}(t, \pi), \overline{x}(t, \pi)]} \gamma^t r(x) = \hat{v^r}(\pi)
\end{equation*}
